---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

# Intro

* [plumber docs: Introduction](https://www.rplumber.io/)

```{r}
install.packages("plumber")
```

```{r}
# plumber.R

#* Echo back the input
#* @param msg The message to echo
#* @get /echo
function(msg=""){
  list(msg = paste0("The message is: '", msg, "'"))
}

#* Plot a histogram
#* @png
#* @get /plot
function(){
  rand <- rnorm(100)
  hist(rand)
}

#* Return the sum of two numbers
#* @param a The first number to add
#* @param b The second number to add
#* @post /sum
function(a, b){
  as.numeric(a) + as.numeric(b)
}

```

Now run it.

```{r}
library(plumber)
r <- plumb("./projects/plumber_test/plumber.R")  # Where 'plumber.R' is the location of the file shown above
r$run(port = 8000)
```

You can visit this URL using a browser or a terminal to run your R function and get the results. For instance `http://localhost:8000/plot` will show you a histogram, and `http://localhost:8000/echo?msg=hello` will echo back the 'hello' message you provided. On `http://127.0.0.1:8000/__swagger__/` the Swagger UI definition will be exposed.

Here we're using curl via a Mac/Linux terminal.

```{bash eval=FALSE}
$ curl "http://localhost:8000/echo"
 {"msg":["The message is: ''"]}
$ curl "http://localhost:8000/echo?msg=hello"
 {"msg":["The message is: 'hello'"]}
```

# Creating APIs in R with Plumber

* [Creating APIs in R with Plumber](https://www.rplumber.io/docs/index.html)


